Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
name introduction

\begin{figure*}[h]
  \centering
  \begin{tabular}{rc}\toprule
    \sc{Task}& \begin{tabular}{rcl}
      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
      Pushmeet Kohli&$\longrightarrow$&???
    \end{tabular}
      \\\midrule
    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
      \\\midrule
    \sc{DSL}& \begin{tabular}{l}
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
      \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
      \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}
      \\\bottomrule
  \end{tabular}
\end{figure*}

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL that provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Starting from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
and iterates between three different steps: an
\textbf{Explore} step uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the DSL by discovering regularities in the programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to three domains:
Lisp-style functions on lists;
FlashFill-style~\cite{gulwani2011automating} string processing; and
symbolic regression;
For each of these we initially provide a generic
set of programming primitives in a Lisp-like language. % TODO: we don't actually give lisp primitives for regression..
% including conditionals, variables, and higher-order functions.
Our algorithm then discovers
its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s

>>>>

Imagine that you are asked to edit some text,
and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
From this example,
you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
drawing upon your prior knowledge of text,
like that words are separated by spaces and that
one commonly prepends titles like ``Dr.''
In this work,
we consider the problem of
building agents that solve few-shot learning tasks like these,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
We think of
solutions to these tasks as being best represented by programs,
and so our problem can be stated as follows:
how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Pushmeet Kohli&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}
We combine these two ideas into
a new algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

We call the learned library of subroutines a \textbf{Domain-Specific
  Language} (\textbf{DSL}), a term widely used in the program
synthesis community~\cite{polozov2015flashmeta}.
Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
The contribution of this work is
an algorithm for learning DSLs while jointly training a neural net to
search for programs in the DSL.

%% synthesize programs for task domains such as string transformations,
%% list processing, and robot navigation and planning. However, all these
%% approaches -- symbolic, neural and neural-symbolic -- rely upon a
%% hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
%% inventory of restricted programming primitives, encoding domain-specific
%% knowledge about the space of programs. In
%% practice
%% we often have only a few input/output examples for each
%% program to be induced, and thus success often hinges on having a good
%% DSL that provides a crucial inductive bias for what would otherwise be an
%% unconstrained search through the space of all computable functions.
%% Here we ask, to what extent can we dispense with such highly
%% hand-engineered domain-specific languages?

%% We propose \emph{learning} the DSL by inducing a library of domain--specific
%% subroutines. We consider the setting where we have a collection of
%% related programming tasks, each specified by a set of input/output
%% examples. %% We do not assume that the tasks are annotated with
%% %% ground-truth programs.
%% Starting from a weaker or more
%% general library of primitives, we give an algorithm for constructing a richer, more powerful,
%% and better-tuned DSL.

Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
because it iterates between three different steps: an
\textbf{Explore} step uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the DSL by discovering regularities in the programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing;  symbolic regression; and turtle graphics~\cite{turtle}.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&\textbf{75\%} &\% &33\%&0\%&30\%\\
  Solve time&  29s&s&80s&--&
